---
title: "Project2"
author: "Holly Cui, Elyse McFalls"
date: "2023-12-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(ggplot2)
```

In our design, there are 16 conditions of paper planes as follows:

ID     Treatment                                ID           Treatment
---    ----------------------                   --------     --------------------
a      Only high rotor length                   bd           High leg length & leg clip
b      Only high leg length                     cd           High leg width & leg clip
c      Only high leg width                      abc          High rotor length & high leg length & high leg width
d      Only leg clip                            abd          High rotor length & high leg length & leg clip
ab     High rotor length & high leg length      acd          High rotor length & high leg width & leg clip
ac     High rotor length & high leg width       bcd          High leg length & high leg width & leg clip
ad     High rotor length & leg clip             abcd         High on all factors
bc     High leg length & high leg width         1            Low on all factors

where the levels of highs and lows are defined by: 

  - **Rotor length**: low = 7.5 cm, high = 8.5 cm
  - **Leg length**: low = 7.5 cm, high = 12.0 cm
  - **Leg width**: low = 3.2 cm, high = 5.0 cm
  - **Leg clip**: no, yes


```{r}
set.seed(123)
trt = c(rep("a", 5), rep("b", 5), rep("c", 5), rep("abc", 5))
```

